---
title: "Assignment 16 Scott Van Slyck"
author: "Scott Van Slyck"
date: "4/21/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Install and load the packages needed for this assignment

```{r echo = FALSE}
# list of packages require for this assignment
packages = c("tidyverse", "ggplot2")

## now load or install and load
package.check <- lapply(
  packages,
  FUN = function(x) {
    if (!require(x, character.only=TRUE)) {
      install.packages(x, dependencies=TRUE)
      library(x, character.only=TRUE)
    }
  }
)
```

# User-Defined Functions

One of the great strengths of R is the user's ability to add functions. In fact, many of the functions in R are actually functions of functions. The structure of a function is given below.

```
function_name <- function(opt1, opt2, ... ){
  # Comments
  code
  return(object)
}

```

Objects in the function are local to the function. The object returned can be any data type.

 - Vector
 - Matrix
 - Data Frame
 - List
 
Let's do a quick review.

## Review of R Data Types

**Question 1.** Create a numeric vector of all the odd integers from 1 to 99.

```{r}
num <- seq(1,99, by = 2)
num
```

**Question 2.** Create a character vector of your two favorite flavors of ice cream.

```{r}
char <- c("butter pecan", "pralines and cream", "cherry")
char
```


**Question 3.** Create a 3 by 3 numeric matrix.

```{r}
mat <- matrix(1:9, nrow = 3, ncol = 3)
mat
```

**Question 4.** Create a list from the objects `num`, `char`, and `mat`.

```{r}
my_list <- list(num, char, mat)
names(my_list) <- c("odd_integers", "ice_cream", "matrix")
my_list
```


## Functions

### 75th Percentile

Now let's write a function that takes in a numeric vector and returns the value of the 75th percentile.

```{r}
get_q75 <- function(n){
  # Purpose
  # -------
  # Return the 75th percentile of a numeric vector
  # 
  # Parameters
  # ----------
  # n : A numeric vector
  q75 <- quantile(n)[4]
  names(q75)[1] <- "75th Percentile"
  return(q75)
}
```

Now we can use our function. But first, we need a vector. The `sample` function takes a sample of the specified size from the elements of a vector of one or more elements using either with or without replacement.

```{r}
sample(1:100, 20)
sample(letters, 26, replace = TRUE)
```
The `runif` function function in R programming language is used to generate a sequence of random following the uniform distribution. 

```{r}
runif(10)
runif(10, -1, 1)
```

```{r}
x <- sample(1:100, 20)
get_q75(x)
```

```{r}
x <- runif(20)
get_q75(x)
```
Suppose we want to make sure that no one tries to use our function on a vector that's not numeric. We can add an `if` statement to make sure the data type of the input vector is numeric.

```{r}
get_q75 <- function(n){
  # Purpose
  # -------
  # Return the 75th percentile of a numeric vector
  # 
  # Parameters
  # ----------
  # n : A numeric vector
  if (!is.numeric(n)){
    return("The vector argument must be of type numeric.")
  } else {
    q75 <- quantile(n)[4]
    names(q75)[1] <- "75th Percentile"
    return(q75)
  }
}
```

This should not work on a vector that is not numeric. Let;s check to be sure.

```{r}
x <- sample(letters, 25, replace = TRUE)
get_q75(x)
```

### `titanic`

```{r}
url <- "https://raw.githubusercontent.com/mahmoudharding/dsc495/main/lectures/data/titanic.csv"
titanic <- read_csv(url)
```


Our function works on a single numeric vector and will return a message if the vector is not numeric. But what if we want to use our function on a column from a data frame. We will need a built-in R function and method to complete our task.


- Given a matrix or data frame n, `t()` returns the transpose of n as a matrix.

- The `as.vector()` is a built-in R method that converts a matrix into a vector.


```{r}
get_q75 <- function(n){
  # Purpose
  # -------
  # Return the 75th percentile of a numeric vector
  # 
  # Parameters
  # ----------
  # n : A numeric vector
  if (is.data.frame(n) | is_tibble(n)){
    n <-t(n)
    n <- as.vector(n)
  }
  if (!is.numeric(n)){
    return("The vector argument must be of type numeric.")
  } else {
    q75 <- quantile(na.omit(n))[4]
    names(q75)[1] <- "75th Percentile"
    return(q75)
  }
}
```


Now we can use our function on a column from a data frame.

```{r}
print("75th percentile of age on the Titanic")
get_q75(titanic %>% select(Age))
print("75th percentile of fare on the Titanic")
get_q75(titanic %>% select(Fare))
```

### The Variance

What if we wanted to return more than one object from our function. We can do that using a list. Let's write a function that returns the sample coefficient of variation along with the mean and standard deviation of a numeric vector.

```{r}
get_coefvar <- function(n){
  # Purpose
  # -------
  # Return the coefficient of variation, mean, and
  # standard deviation of a numeric vector
  # 
  # Parameters
  # ----------
  # n : A numeric vector
  if (is.data.frame(n) | is_tibble(n)){
    n <-t(n)
    n <- as.vector(n)
  }
  if (!is.vector(n, mode = "numeric")){
    return("The vector argument must be of type numeric.")
  } else {
    mean <- mean(n, na.rm = TRUE)
    stdev <- sd(n, na.rm = TRUE)
    cov <- mean/abs(stdev)
    return(list(mean = mean, stdev = stdev, coef_var = cov))
  }
}

```

```{r}
get_coefvar(titanic %>% select(Age))
```
## Simple Linear Regression

Simple linear regression is used to predict a quantitative outcome $y$ on the basis of one single predictor variable $x$. The goal is to build a mathematical model (or formula) that defines $y$ as a function of the $x$ variable.

```{r}
mtcars %>% head()
```

```{r}
g <- ggplot(data = mtcars, aes(x = wt, y = mpg))
g + geom_point()
```

```{r}
g + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

The graph above suggests a linearly decreasing relationship between the weight and the mpg variables.

It’s also possible to compute the correlation coefficient between the two variables using the R function `cor()`.

```{r}
cor(mtcars$wt, mtcars$mpg)
```

Simple linear regression tries to find the best line to predict mpg on the basis of the weight of the vehicle. The linear model equation can be written as $\text{mpg } = b_0 + b_1 \cdot \text{ weight}$. The R function `lm()` can be used to determine the beta coefficients of the linear model.

```{r}
mpg_model <- lm(mpg  ~ wt, data = mtcars)
mpg_model
```

The equation for our linear model is $\hat{\text{mpg }} = 37.285 - 5.344 \cdot \text{ weight}$. To access the coefficients we can look at the structure of the `mpg_model` object.

```{r}
str(mpg_model)
```

Since this is a named list we can access the elements using the name. For example, if we want the `coefficients` we can type `mpg_model$coefficients`.

```{r}
mpg_model$coefficients
paste("The y-intercept is", mpg_model$coefficients[1])
paste("The slope is", mpg_model$coefficients [2])
```

We can use these values to overlay the graph of the line on the scatter plot.

```{r}
g + geom_point() + geom_abline(intercept = mpg_model$coefficients[1],
                     slope = mpg_model$coefficients[2], color = "red")
```
The `mpg_model` object also contains summary information.

```{r}
summary(mpg_model)
```

## Exercises

**Exercise 1.** In the code chunk below, write a function that will return the standardized values in a numeric vector.

**Hint:** The formula to standardize a vector $z_i = \frac{x_i-\text{ mean of data}}{\text{standard deviation of data}}$

```{r}
standard_units <- function(n){
  mean = mean(n, na.rm = TRUE)
  stdev = sd(n, na.rm = TRUE)
  return (n - mean)/stdev
}
```

Check your function by running the code chunk below.

```{r}
standard_units(titanic$Age)
```


**Exercise 2.** During class we mde a scatter plot to visualize the relationship between `wt` (weight) and `mpg` (miles per gallon). In the code chunk below, make a scatter plot of the relationship between `hp` and `mpg`. Overlay the scatter plot with the line of besst fit.

```{r}
g <- ggplot(data = mtcars, aes(x = hp, y = mpg))
g + geom_point() + geom_smooth(method = "lm", se = FALSE)
```

**Exercise 3.** In the code chunk below, fit a linear model using `hp` as the predictor $(x)$ and `mpg` as the response $(y)$. Display the summary information of the model object.


```{r}
hp_model <- lm(mpg ~ hp, data = mtcars)
summary(hp_model)
```